Learning Auditory Similarities Between Instrumental Playing Techniques

### 2020-05-19 paper/paper.tex
improve scattering section

% enjoys a thirty-fold reduction in dimensionality, while covering a  time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
%We refer to \cite{bellet2015book} for a review of the state of the art in metric learning.

@@

Let $\boldsymbol{\phi}_T$ be a Gaussian low-pass filter.
We define the joint time--frequency scattering coefficients of the signal $\boldsymbol{x}$ as the four-way tensor
\begin{align}
\mathbf{S_2}(t, \lambda, \alpha, \beta)
=&
\mathbf{U_2}\boldsymbol{x}
\circledast
\big(\boldsymbol{\phi}_T  \otimes \boldsymbol{\phi}_F\big)
(t, \lambda)
\nonumber \\
=&
\iint_{\mathbb{R}^2}
\mathbf{U_1}\boldsymbol{x}(t - t^\prime, \log \lambda - u^\prime, \alpha, \beta)
\,\boldsymbol{\phi}_T (t^\prime)
\,\boldsymbol{\phi}_F (u^\prime)
\;\mathrm{d}t^\prime \, \mathrm{d}u^\prime,
\end{align}
where the symbol $\otimes$ denotes an outer product between two vectors.
In the equation above, the sample rate of time $t$ is proportional to $T^{-1}$ and the sample rate of log-frequency $u=\log_2 \lambda$ is proportional to $F^{-1}$.
Furthermore, the rate $\alpha$ spans along a geometric progression ranging from $T^{-1}$ to $\lambda/Q$.
In the following, we set the time constant to $T=\SI{1000}{\milli\second}$ unless specified otherwise.

The tensor $\mathbf{S_2}$ bears a strong resemblance with the idealized response of a spectrotemporal receptive field (STRF) at the rate $\alpha$ and the scale $\beta$.
Nevertheless, in comparison with the representation known as ``full cortical model'' \cite{patil2012ploscompbiol}, joint time--frequency scattering enjoys a thirty-fold reduction in dimensionality while covering a time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
This is due to the multirate discretization scheme applied throughout the application of wavelet convolutions and pointwise modulus nonlinearities.
We refer to \cite{anden2019tsp} for more details on the implementation of joint time--frequency scattering.

@@

@@

Let us recall the full definition of joint time--frequency scattering (see \lnameref{sec:methods} section):
\begin{equation}
\mathbf{S_2}\boldsymbol{x}(t,\lambda,\alpha,\beta) = 
\left(
\Big\vert
\big\vert
\boldsymbol{x} \overset{t}{\ast} \boldsymbol{\psi}_{\lambda}
\big\vert
\circledast
\boldsymbol{\Psi}_{\alpha,\beta}
\Big\vert
\circledast
\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\phi}_{F})
\right)(t,\lambda),
\label{eq:joint-scattering}
\end{equation}
where the $\boldsymbol{\psi}_{\lambda}$ denotes a Morlet wavelet of center frequency $\lambda$ and resp. $\boldsymbol{\phi}_T$ denotes a Gaussian low-pass filter of width $T$.
Besides the \emph{joint} time--frequency scattering transform, the generative grammar of scattering transforms \cite{lostanlen2019chapter} also encompasses the \emph{separable} time--frequency scattering transform:
\begin{equation}
\mathbf{S_2^{\mathrm{sep}}}\boldsymbol{x}(t,\lambda,\alpha,\beta) = 
\left(
\bigg\vert
\Big\vert
\big\vert
\boldsymbol{x} \overset{t}{\ast} \boldsymbol{\psi}_{\lambda}
\big\vert
\ast
\boldsymbol{\psi}_{\alpha}
\Big\vert
\circledast
\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\psi}_{\beta}\big)
\bigg\vert
\circledast
\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\phi}_F\big)
\right)(t, \lambda),
\label{eq:separable-scattering}
\end{equation}
which has been applied successfully to classify environmental sounds \cite{bauge2013icassp} as well as playing techniques involving periodic modulations \cite{wang2019ismir}.

>>>>

@@

Let $\boldsymbol{\phi}_T$ be a Gaussian low-pass filter.
We define the joint time--frequency scattering coefficients of the signal $\boldsymbol{x}$ as the four-way tensor
\begin{align}
\mathbf{S_2}(t, \lambda, \alpha, \beta)
=&
\mathbf{U_2}\boldsymbol{x}
\circledast
\big(\boldsymbol{\phi}_T  \otimes \boldsymbol{\phi}_F\big)
(t, \lambda)
\nonumber \\
=&
\iint_{\mathbb{R}^2}
\mathbf{U_1}\boldsymbol{x}(t - t^\prime, \log \lambda - u^\prime, \alpha, \beta)
\,\boldsymbol{\phi}_T (t^\prime)
\,\boldsymbol{\phi}_F (u^\prime)
\;\mathrm{d}t^\prime \, \mathrm{d}u^\prime,
\label{eq:S2}
\end{align}
where the symbol $\otimes$ denotes an outer product between two vectors.
In the equation above, the sample rate of time $t$ is proportional to $T^{-1}$ and the sample rate of log-frequency $u=\log_2 \lambda$ is proportional to $F^{-1}$.
Furthermore, the rate $\alpha$ spans along a geometric progression ranging from $T^{-1}$ to $\lambda/Q$.
In the following, we set the time constant to $T=\SI{1000}{\milli\second}$ unless specified otherwise.

The tensor $\mathbf{S_2}$ bears a strong resemblance with the idealized response of a spectrotemporal receptive field (STRF) at the rate $\alpha$ and the scale $\beta$.
Nevertheless, in comparison with the representation known as ``full cortical model'' \cite{patil2012ploscompbiol}, joint time--frequency scattering enjoys a thirty-fold reduction in dimensionality while covering a time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
This is due to the multirate discretization scheme applied throughout the application of wavelet convolutions and pointwise modulus nonlinearities.

In addition to second-order scattering coefficients (Equation \ref{eq:S2}), we compute joint time--frequency scattering at the first order by convolving the scalogram $\mathbf{U_1 \boldsymbol{x}}$ (Equation \ref{eq:U1}) with the low-pass filter $\boldsymbol{\phi}_T$ over the time dimension, with wavelets $\boldsymbol{\psi}_{\beta}$ ($\beta \geq0$) over the log-frequency dimension, and by applying the complex modulus:
\begin{align}
\mathbf{S_1}\boldsymbol{x}(t, \lambda, \alpha=0, \beta) &=
\big\vert
\mathbf{U_1}\boldsymbol{x} \circledast
\big(\boldsymbol{\phi}_T \otimes \boldsymbol{\psi}_{\beta} \big)
\big\vert (t, \lambda)
\nonumber \\
&=
\left \vert
\iint_{\mathbb{R}^2}
\mathbf{U_1}\boldsymbol{x}(t - t^\prime, \log \lambda - u^\prime)
\,\boldsymbol{\phi}_T (t^\prime)
\,\boldsymbol{\psi}_{\beta} (u^\prime)
\;\mathrm{d}t^\prime \, \mathrm{d}u^\prime
\right \vert,
\label{eq:S1}
\end{align}
where the time constant $T$ is the same as in Equation $\ref{eq:S2}$, \ie{} $T=\SI{1000}{\milli\second}$ by default.
Over the time variable, we set the modulation rate $\alpha$ of $\mathbf{S_1}$ to zero in the equation above.
Conversely, over the log-frequency variable, the edge case $\beta=0$ corresponds to replacing the wavelet $\boldsymbol{\psi}_{\beta}$ by the low-pass filter $\boldsymbol{\psi}_F$, comparably to Equation $\ref{eq:Psi}$.
Lastly, we concatenate 
\begin{equation}
    \mathbf{S}\boldsymbol{x}(t, p) 
\end{equation}
We refer to \cite{anden2019tsp} for more details on the implementation of joint time--frequency scattering.

@@

%We refer to \cite{bellet2015book} for a review of the state of the art in metric learning.

@@

Let us recall the full definition of joint time--frequency scattering (see \lnameref{sec:methods} section):
\begin{equation}
\mathbf{S_2}\boldsymbol{x}(t,\lambda,\alpha,\beta) = 
\left(
\Big\vert
\big\vert
\boldsymbol{x} \overset{t}{\ast} \boldsymbol{\psi}_{\lambda}
\big\vert
\circledast
\boldsymbol{\Psi}_{\alpha,\beta}
\Big\vert
\circledast
\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\phi}_{F})
\right)(t,\lambda),
\label{eq:joint-scattering}
\end{equation}
where the $\boldsymbol{\psi}_{\lambda}$ denotes a Morlet wavelet of center frequency $\lambda$ and resp. $\boldsymbol{\phi}_T$ denotes a Gaussian low-pass filter of width $T$.
Besides the \emph{joint} time--frequency scattering transform, the generative grammar of scattering transforms \cite{lostanlen2019chapter} also encompasses the \emph{separable} time--frequency scattering transform:
\begin{equation}
\mathbf{S_2^{\mathrm{sep}}}\boldsymbol{x}(t,\lambda,\alpha,\beta) = 
\left(
\bigg\vert
\Big\vert
\big\vert
\boldsymbol{x} \overset{t}{\ast} \boldsymbol{\psi}_{\lambda}
\big\vert
\ast
\boldsymbol{\psi}_{\alpha}
\Big\vert
\circledast
\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\psi}_{\beta}\big)
\bigg\vert
\circledast
\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\phi}_F\big)
\right)(t, \lambda),
\label{eq:separable-scattering}
\end{equation}
where the wavelet $\boldsymbol{\psi}_{\lambda}$ has a quality factor of $Q=12$ whereas the wavelets $\boldsymbol{\psi}_{\alpha}$ and $\boldsymbol{\psi}_{\beta}$ have a quality factor of one.
Previous publications have successfully applied separable time--frequency scattering in order to classify environmental sounds \cite{bauge2013icassp} as well as playing techniques involving periodic modulations \cite{wang2019ismir}.
